---
title: "03 - examine and manipulate data"
author: "Jenny Bigman"
date: "5/12/2022"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(knitr, tidyverse, here)

dat <- read_csv(here("./data/portal_data_joined.csv"))
```

Let's work with our data from the last lesson and the dplyr package. We will examine, manipulate, summarize, and plot the data.

[dpylr](https://dplyr.tidyverse.org/), in my opinion, is one of the most useful packages in R. Check out the website and read about it. These R documentation websites usually have very helpful tutorials (vignettes). There is also a helpful blog post by a Sean Anderson (a former lab-mate of mine, https://seananderson.ca/2014/09/13/dplyr-intro/). I used a lot of Sean's examples for this lesson and the mixed effects modeling lesson, so thank you Sean!

dplyr is in the tidyverse, so we don't need to load it separately.

We've already used some dplyr functions! We have used `glimpse()`, `mutate()`, and `%>%` (pipe). Let's build on these. 

dplyr has 5 main functions: `mutate()`, `select()`, `filter()`, `summarize()`, & `arrange()`. 

# Mutate()

We've already worked with `mutate()` but let's examine it more closely. `mutate()` lets you add new columns, which can build on each other if you want.

```{r}

mutate(dat, log_hindfoot_length = log10(hindfoot_length))

# we didn't save this as an object, so this is why the results printed
# out for us

# and if we run names(dat), you won't see your new column
names(dat)

# we can add a new column and write over our dataframe so the column
# will be added for us

dat <- mutate(dat, log_hindfoot_length = log10(hindfoot_length))
names(dat)

# another way to do this (which you have seen before) is to use a %>% (pipe)

# this is really helpful for when you are applying many functions to a 
# dataframe and don't want to keep retyping it

dat <- dat %>%
	mutate(log_hindfoot_length = log10(hindfoot_length))

names(dat)

# this looks like it worked, but let's check the range of values
range(dat$hindfoot_length)
range(dat$log_hindfoot_length)

# any one know why this is NA?

```

Let's remove NAs in our dataset to help us with some later wrangling.
```{r}

# we can create new dataframe with no NAs if we'd like (but keep the original!)
dat_no_NA <- na.omit(dat)

# other way to deal with NAs: complete_cases()
# when entering data into excel to later read in R, 
#leave cell blank and this will be an NA

# don't write NA in the cell!

# now let's try again
dat_no_NA <- dat_no_NA %>%
	mutate(log_hindfoot_length = log10(hindfoot_length))

names(dat_no_NA)

# this looks like it worked, but let's check the range of values
range(dat_no_NA$hindfoot_length)
range(dat_no_NA$log_hindfoot_length)
```

# Select()

Let's work next with `select()`. `select()` allows you to select certain columns for subsetting. 

```{r}
species <- select(dat, species)
# you might want to use this when you have many columns in your raw data that you 
# don't need for analysis -- but you should never change your raw data!

# you can also use '-' in front of the column to remove it
species <- select(dat, -species) 
# this will remove the column from the dataframe
# instead of keeping just this column like the above example

# can use helper functions 
select(dat, contains("i"))
select(dat, starts_with("ta"))
select(dat, ends_with("us"))

# can also use column positions (more like a traditional subset in base R)
select(dat, 1:5) # first five columns

# can also do this in base R (not using the dplyr package)
dat[, 1:5]

# how would you tell if the above produces the same result?
```

Challenge: Can you write the above code 'species <- select(dat, species)' using pipes in the code chunk below? When finished, place your blue sticky note on your laptop. If you need help, put up your pink sticky. 

```{r}


```

```{r}
species <- dat %>%
	select(species)
```

# Filter()

Next up is `filter()`. `filter()` subsets rows by name. This is a cool and useful function because you can use any valid statement. You will see what I mean below.

```{r}

filter(dat, hindfoot_length < 20)

# remember to save this as an object if you want to 

dat_filtered <- filter(dat, hindfoot_length < 20)

# or

dat_filtered <- dat %>% 
	filter(hindfoot_length < 20)

# can also filter many other ways

filter(dat, hindfoot_length == 20)

filter(dat, hindfoot_length >= 20)

# can filter by more than one variable 

filter(dat, species == "fuscus" & weight > 100)

# bonus:

# slice() works similarly to filter but subsets rows by position

slice(dat, 100:101) 

# can also do this in base R and using filter but have to change the code a bit

dat[1:5, ]

# the code to subset columns was dat[ , 1:5].

# What is different?

# using filter
filter(dat)[100:101, ] 

# can also filter for multiple values at once

# what are the years in the dataframe?
unique(dat$year)

# let's filter out anything before 1980
years_to_keep <- 1980:2002

dat_trim <- dat %>%
	filter(year %in% years_to_keep)

```

Challenge: How many rows are in a new dataset that only includes individuals of the harrisi species that have a hindfoot length 

```{r}

```

```{r}

nrow(filter(dat, species == "harrisi" & hindfoot_length < 50))
```


# Summarize()

Next, let's look at `summarize()`. This is the same as `summarise()` [Hadley Wickham, developer of the tidyverse, is from New Zealand]. `summarize()` allows you to, well, summarize data. This function is especially useful when combined with another function, `group_by()`, which allows you to group data and then summarize. 

```{r}

# summarize by itself

sum <- dat %>% 
	summarise(mean_hindfoot_length = mean(hindfoot_length))

```

Check the new object that we just created. Why does it NA? 

We need to specify that we have NAs in our dataset, but we don't care about them for the purpose of summarizing. 

```{r}

sum <- summarise(dat, mean_hindfoot_length = mean(hindfoot_length, na.rm = TRUE))

```

Challenge: How might we write the above code in another way (hint: use the pipe [%>%])

```{r}

```


```{r}
sum <- dat %>%
	summarize(mean_hindfoot_length = mean(hindfoot_length, na.rm = TRUE))
```

Great, now we have a mean hindfoot length. But there are other ways to do this outside of summarize and we haven't fully unleashed the power of summarize. 

First, let's look at a simple way to calculate the mean of a column (or vector).

```{r}

mean(dat$hindfoot_length, na.rm = TRUE) # gives us the same answer

# can also do range(), min(), max(), median(), etc.

# check out the stats package documentation:
# https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html

# you can also subset a vector, but not using dplyr's filter() or select(), which are for dataframes
vector <- dat$hindfoot_length

# subset the hindfoot lengths > 50
vector[vector > 50]

# subset the hindfoot length > 50 but < 100
vector[vector > 50 & vector < 100]

# what about the hindfoot lengths less < 20 but > 100
vector[vector < 20 | vector > 100]

```


Now let's unleash the power of using `group_by()` and `summarize()` together! I use `group_by()` all of the time. It is a really useful function, and not just with summarize.

```{r}

# count the number of individuals per species
count_sum <- dat %>%
	group_by(species) %>%
	summarize(n())

# mean hindfoot length by species
mean_hfl <- dat %>%
	group_by(species) %>%
	summarize(mean_hindfoot_length = mean(hindfoot_length, na.rm = TRUE))

# group_by plot_id -- how many records per plot?

plot_id_count <- dat %>%
	group_by(plot_id) %>%
	summarise(n())

```

# Arrange()

Last, let's look at `arrange()`. `arrange()` let's you order the rows in ascending or descending order by a particular column.

```{r}

arrange(dat, hindfoot_length)

arrange(dat, year, hindfoot_length) # can specify more than one column!

arrange(dat, desc(hindfoot_length))
```

Challenge: What if I wanted to only see the first five rows and arrange by `record_id` in descending order?
```{r}

```

```{r}

head(arrange(dat, desc(record_id)))
		 
``` 


There are other functions in dplyr. There is even a whole philosophy on data wrangling using dplyr -- `split - apply - combine`. We are doing a bit of this already with `group_by()`. The apply family are another set of functions that are very useful, but a little more advanced (they are especially useful when writing your own functions -- which yes! you can write your own). We won't go into them now but good to be aware of them. If you are interested, here are some resources for dplyr and data wrangling in general:


1. [dplyr vignette](https://dplyr.tidyverse.org/articles/dplyr.html)
2. [grouping](https://dplyr.tidyverse.org/articles/grouping.html)
3. [joining dataframes using dplyr](https://dplyr.tidyverse.org/articles/two-table.html)
4. [wrangling columns](https://dplyr.tidyverse.org/articles/colwise.html)
5. [wrangling rows](https://dplyr.tidyverse.org/articles/rowwise.html)


Before we plot some data, let's look another useful function from dplyr, `rename()`. `rename()` allows you to rename column headings. 

```{r}

names(dat)

dat <- dat %>%
	rename(species_name = species) # I want to rename species to species_name

```

 Ok, let's move on to `ggplot()`.